---
title: "Homework 3"
author: "Drew Dahlquist"
date: "2/17/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.a)

$$\\[3in]$$

Expert 2's prior is uninformative since solving for the desired mean and standard deviation yiels a = b = 17.5. 

1.b)

The posterior for Expert 1 is relatively unchanged given all 5 mice test positive, since it expected 0.95 probability for each mouse, whereas the posterior for Expert 2 has a greater mean and lower variance.

```{r 1b, echo=FALSE}
# create a set of values between 0 and 1 to compute the density at
theta = seq(0,1, length=100)

# data
Y = 5
n = 5

# prior 1
a.1 = 17.1
b.1 = 0.9

# prior 2
a.2 = 17.5
b.2 = 17.5

# posterior 1
A.1 = Y + a.1
B.1 = n - Y + b.1

# posterior 2
A.2 = Y + a.2
B.2 = n - Y + b.2

# posterior 1
posterior.1 = dbeta(theta, A.1, B.1)

# posterior 2
posterior.2 = dbeta(theta, A.2, B.2)

# plotting
plot(theta, posterior.1, main="Posteriors under each expert's prior", ylab="density", type="l", col="2", ylim=c(0,6)) # posterior 1
lines(theta, posterior.2, type="l", col="4") # posterior 2

# add legend
legend("topleft", c("Expert 1", "Expert 2"), lty=c(1,1),col=c(2,4))
```

2.a)

The Negative Binomial distribution is used to count the number of failures from independent trails with probability of success $\theta$ before recording the mth success.

2.b)

$$\\[1in]$$

2.c)

```{r 2c, echo=FALSE}
m = 5
Y = 10
a = b = 1

theta = seq(0,1, length=100)
plot(theta, dbeta(theta, a+m, b+Y), main="Posterior when m = 5, Y = 10, a = b = 1", ylab="density", type="l", col="2", ylim=c(0,6))
```

    95% credible interval = (`r qbeta(c(.025,.975),a+m,b+Y)`).

3)

$$\\[4in]$$

4.a)

$$\\[.5in]$$

```{r 4a, echo=FALSE}
lambda = seq(0,5,length=500)

plot(lambda, 1/sqrt(lambda), main="Plot of Jeffreys' Prior for lambda ~ Poisson", ylab="Jeffreys' Prior", type="l", col="4")
```

4.b)

No, this is not a proper prior since the integral does not converge.

4.c)

$$\\[1in]$$
